---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

##Loading and preprocessing the data
This is a R markdown file. We input activity.csv to data.  

```{r, echo = TRUE}
data <- read.csv("activity.csv")
data$steps <- as.numeric(data$steps)
head(data)
```

We then create a new data frame with total number of steps taken in a day.
```{r, echo = TRUE}
dataProcessed <- aggregate(data$steps, by=list(date=data$date), FUN=sum)
colnames(dataProcessed)<-c("date", "steps")
dataProcessed$steps <- as.numeric(dataProcessed$steps)
```


##What is mean total number of steps taken per day?

We produce a histogram showing the total number of steps taken each day. 
```{r, echo = TRUE}
hist(dataProcessed$steps, col = "red", xlab = "Steps taken in a day", main = "Histogram of steps taken in a day")
mea <- mean(dataProcessed$steps)
med <- median(dataProcessed$steps)
```

The mean number of steps taken in a day is `r mea` and the median is `r med`

##What is the average daily activity pattern?
```{r, echo = TRUE}
dataUse <- data
dataUse <- dataUse[complete.cases(dataUse),]
dataInterval <- aggregate.data.frame(dataUse$steps, by=list(dataUse$interval), FUN=mean)
colnames(dataInterval)<-c("interval", "mean_interval_steps")
plot(dataInterval$interval, dataInterval$mean_interval_steps, type = "l", ylab = "Steps", xlab = "Interval")
maxInterval <- dataInterval[which.max(dataInterval$mean_interval_steps), 1]
```

The Interval with the highest number of steps is `r maxInterval`.

##Imputing missing values
Calculate number of NAs.
```{r, echo = TRUE}
sum(is.na(data$steps))
```

Create a new data frame with NAs replaced with interval's mean and plot.
```{r, echo = TRUE}
dataNAremove <- data
dataNAremove <- merge(dataNAremove, dataInterval, by = "interval")
dataNAremove$steps[ is.na(dataNAremove$steps) ] <- dataNAremove$mean_interval_steps
head(dataNAremove)
dataProcessed2 <- aggregate(dataNAremove$steps, by=list(date=dataNAremove$date), FUN=sum)
colnames(dataProcessed2)<-c("date", "steps")
dataProcessed2$steps <- as.numeric(dataProcessed2$steps)
hist(dataProcessed2$steps, col = "red", xlab = "Steps taken in a day", main = "Histogram of steps taken in a day")
mea <- mean(dataProcessed2$steps)
med <- median(dataProcessed2$steps)
```

There is very little difference between histograms.  
The new mean number of steps taken in a day is `r mea` and the new median is `r med`


##Are there differences in activity patterns between weekdays and weekends?

Create factors and plot based on factors.
```{r, echo = TRUE}
weekends <- c('Saturday', 'Sunday')
dataNAremove$date <- as.Date(dataNAremove$date)
dataNAremove$weekend <- factor((weekdays(dataNAremove$date) %in% weekends), levels=c(TRUE, FALSE), labels=c('weekend', 'weekday'))
dataWeekday <- subset(dataNAremove, weekend == 'weekday')
dataWeekend <- subset(dataNAremove, weekend == 'weekend')
par(mfrow = c(2,1))
dataInterval <- aggregate.data.frame(dataWeekend$steps, by=list(dataWeekend$interval), FUN=mean)
colnames(dataInterval)<-c("interval", "mean_interval_steps")
plot(dataInterval$interval, dataInterval$mean_interval_steps, type = "l", ylab = "Steps", xlab = "Interval", main = "Mean steps over 5-minute intervals on the weekend")
dataInterval <- aggregate.data.frame(dataWeekday$steps, by=list(dataWeekday$interval), FUN=mean)
colnames(dataInterval)<-c("interval", "mean_interval_steps")
plot(dataInterval$interval, dataInterval$mean_interval_steps, type = "l", ylab = "Steps", xlab = "Interval", main = "Mean steps over 5-minute intervals on the weekday")
```


